Realizing Temporal Graphs

### 2023-01-30 RealizingTemporalGraphs.tex
Monday 30th Jan

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Observe also that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments.
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

    If $S_{u,v}$ is not of length $4$, we fix an arbitrary vertex $v_i \in S_{u,v}$.

@@

@@

\subsection{Adding constraints and variables to the ILP}
First case we observe is the case when we want to determine the labels on fastest temporal paths among vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P$, with respect to the label of the first edge (see \cref{claim:FPT-uv-Labelalledges}).
%Note that since we determined labels of all edges, with respect to the first edge, we also know the exact values of waiting times at each inner vertex of $P_{u,v}$, when traversing the path.
In the case when $P_{u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{claim:FPT-uv-LabelAlmostalledges}). \todo{Chenge/explain}
%In both cases we introduce the equality constraint for $P_{u,v}$ (as in \cref{eq:FPT-equalityConstraint}).
%Since there are $\ell - 1$ edge labels that are not determined with respect to the label of the first edge, may be some  TODO from above??

@@

\begin{claim}
    After determining the labels of fastest temporal paths between all pairs of vertices of interest,
    there can be at most one not yet determined segment among each pair of them.
    %among any pair of vertices of interest $u,v \in U$ there can be at most one undetermined segment $S = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
   Suppose for the contradiction that after determining the labels of all fastest temporal paths between vertices of interest,
   there are two not yet determined segments among $u,v \in U$.
   Denote them with $S_{u,v}$ and $S'_{u,v}$. 
   Then $C = S_{u,v} \cup S_{u,v}'$ forms a cycle in $G'$. Since $G'$ admits no loops, the cycle $C$ is of length at least $3$.
   Since $C$ is a cycle, at least one of its edges $e \in E(C)$ has to be in the feedback edge set $F$, and therefore, by the definition of vertices of interest $U$, 
   both endpoints of $e$ have to be in $U$.
   If $e = uv$ then there are only two vertices of interest in $C$, but the label of edge $e$ is determined.
   And if $e \neq uv$ then there are at least $3$ vertices of interest in $C$.
   So $S_{u,v}$ and $S_{u,v}'$ cannot be both fully not yet determined.
\end{proof}

%From the above it follows that there are at most $O(|U|^2) = O(k^2)$ undetermined segments in $G'$ after the algorithm finishes with determining the labels for all fastest temporal paths among vertices of interest.

@@

        all fastest temporal paths from $x$ to any vertex of interest $u'$ go through the edge $z_{i-1}x$ (not trough $p$).
        Now, if the edges of tree are determined with respect to the label of the edge $z_{i-1}x$,

>>>>

For an easier understanding and the analysis of the algorithm we give the following definition.
\begin{definition}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
    Note $S_{u,v} \neq S_{v,u}$.
\end{definition}
%From the definition it follows, that there are $O(k^2)$ segments in $G'$. \todo{True but does not follow directly just from the definition}
Observe that a temporal path in $G'$ between two vertices of interest is either a segment, or consists of a sequence of some segments.
In the analysis of the algorithm we rely heavily on the fact that there are $O(k^2)$ segments in $G'$, 
which we prove later on (see~\cref{obs:FPT-k2segments}).
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

@@

    If $S_{u,v}$ is not of length $4$, we fix an arbitrary vertex $v_i \in S_{u,v} \setminus \{u,v\}$.

@@

    Note that this case results also in knowing the fastest paths from the vertex $v_i \in S_{u,v}$ to $w,z \in S_{w,z}$ for all segments $S_{w,z}$, 
    \ie we know the fastest paths from a fixed $v_i \in S_{u,v}$ to all vertices of interest in $U$.

@@

\subsection{Adding constraints and variables to the ILP}
First case we observe is the case when we want to determine the labels on fastest temporal paths among vertices of interest.
We proceed in the following way.
Let $u,v \in U$ be two vertices of interest and let $P_{u,v}$ be the fastest temporal path from $u$ to $v$.
If $P_{u,v}$ is a segment we determine all the labels of edges of $P$, with respect to the label of the first edge (see \cref{claim:FPT-uv-Labelalledges}).
%Note that since we determined labels of all edges, with respect to the first edge, we also know the exact values of waiting times at each inner vertex of $P_{u,v}$, when traversing the path.
In the case when $P_{u,v}$ is a sequence of $\ell$ segments, we determine all but $\ell - 1$ labels of edges of $P_{u,v}$, with respect to the label of the first edge (see \cref{claim:FPT-uv-LabelAlmostalledges}). \todo{Change/explain what ``with respect to first label'' means}
%In both cases we introduce the equality constraint for $P_{u,v}$ (as in \cref{eq:FPT-equalityConstraint}).
%Since there are $\ell - 1$ edge labels that are not determined with respect to the label of the first edge, may be some  TODO from above??

@@

\begin{claim}
    After determining the labels of fastest temporal paths between all pairs of vertices of interest,
    there can be at most one not yet determined segment among each pair of them.
    %among any pair of vertices of interest $u,v \in U$ there can be at most one undetermined segment $S = (u, v_1, v_2, \dots, v_p, v)$, where $p \geq 0$ and $v_1,v_2,\dots,v_p \in V(G') \setminus U$.
\end{claim}
\begin{proof}
   Suppose for the contradiction that after determining the labels of all fastest temporal paths between vertices of interest,
   there are two not yet determined segments among $u,v \in U$.
   Denote them with $S_{u,v}$ and $S'_{u,v}$. 
   Then $C = S_{u,v} \cup S_{u,v}'$ forms a cycle in $G'$. Since $G'$ admits no loops, the cycle $C$ is of length at least $3$.
   Since $C$ is a cycle, at least one of its edges $e \in E(C)$ has to be in the feedback edge set $F$, and therefore, by the definition of vertices of interest $U$, 
   both endpoints of $e$ have to be in $U$.
   If $e = uv$ then there are only two vertices of interest in $C$, but the label of edge $e$ is determined.
   And if $e \neq uv$ then there are at least $3$ vertices of interest in $C$.
   So $S_{u,v}$ and $S_{u,v}'$ cannot be both fully not yet determined.
\end{proof}
From the above it follows that
after the algorithm finishes with determining the labels for all fastest temporal paths among vertices of interest
there is at most one fully determined segment among two vertices of interest and at most one undetermined segment.
Therefore the following holds.
\begin{observation}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{observation}

@@

        all fastest temporal paths from $x$ to any vertex of interest $u'$ go through the edge $z_{i-1}x$, not trough $p$ 
        (this is true as if a fastest temporal path from $x$ to some vertex of interest $w'$ went through $p$, then $p$ would be determined).
        Now, if the edges of tree are determined with respect to the label of the edge $z_{i-1}x$ (not $p$),
